---
title: "MVC_Mass"
author: "Shriram Karthikeyan"
date: "November 8, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Delete the column TRVL_DIRC_DESCR from the 2019 and 2019 file
Replace the first row for all the csv files below from the file Data/MassDOT_IMPACT_Template_Fields_Person.csv
```{r}
PLCD_2019 <- read.csv("Data/2019_Person_Level_Crash_Details_.csv", stringsAsFactors=FALSE)
PLCD_2018 <- read.csv("Data/2018_Person_Level_Crash_Details_.csv", stringsAsFactors=FALSE)
PLCD_2017 <- read.csv("Data/2017_Person_Level_Crash_Details_.csv", stringsAsFactors=FALSE)
PLCD_2016 <- read.csv("Data/2016_Person_Level Crash_Details .csv", stringsAsFactors=FALSE)
PLCD_2015 <- read.csv("Data/2015_Person_Level_Crash_Details_.csv", stringsAsFactors=FALSE)
#PLCD_2014 <- read.csv("Data/2014_Person_Level_Crash_Details_.csv", stringsAsFactors=FALSE)

```


```{r}
# PLCD_2019 <- PLCD_2019[ , -which(names(PLCD_2019) %in% c("TRVL_DIRC_DESCR"))]
# PLCD_2018 <- PLCD_2018[ , -which(names(PLCD_2018) %in% c("TRVL_DIRC_DESCR"))]

```


```{r}
PLCD<- rbind(PLCD_2015,PLCD_2016,PLCD_2017,PLCD_2018,PLCD_2019)
#rm(PLCD_2014)
rm(PLCD_2015)
rm(PLCD_2016)
rm(PLCD_2017)
rm(PLCD_2018)
rm(PLCD_2019)
```



add libraries
```{r}
library(dplyr)
library(magrittr)
library(stringr)
library(lubridate)
library(ggplot2)
library(tidyr)
```

```{r}
#summary(PLCD)
na_count<-sapply(PLCD, function(x) sum(is.na(x)))
sort(na_count,decreasing = TRUE)
```




```{r}



PLCD <- PLCD[ , -which(names(PLCD) %in% c("AADT_Station_Number.linked_RD","Milemarker","Opposing_Direction_Speed_Limit.linked_RD","Undivided_Left_Shoulder_Width.linked_RD","Number_of_Peak_Hour_Lanes.linked_RD","AADT_Derivation.linked_RD","Undivided_Left_Shoulder_Type.linked_RD","Truck_Exclusion_Type.linked_RD","Truck_Exclusion_Time.linked_RD","Left_Shoulder_Width.linked_RD ","Median_Width.linked_RD","Vehicle_Sequence_of_Events_.All_Vehicles.","Vehicle_Travel_Direction_.All_Vehicles.","Distance_and_Direction_from_Intersection","Distance_and_Direction_from_Milemarker","Vehicle_Configuration_.All_Vehicles.","RPA","Is_Geocoded","Geocoding_Method","Vehicle_Owner_City_Town"))]

#write.csv(PLCD,'PLCD.csv')

```



```{r}
summary(PLCD$Driver_Age)
hist(PLCD$Driver_Age,nclass = 100)


PLCD$DRIVER_DISTRACTED_TYPE_DESCR
  
# PLCD %>%
#   #select(DRIVER_AGE)%>%
#   filter(DRIVER_AGE<5)
```


Fatal injury analysis:
where did people die and how many?
what was the time of the fatal injuries?
what was the age distribution of the fatal injuries
```{r}
# PLCD%>%
#   filter(NUMB_FATAL_INJR>0)
```



What is the trend of crashes affected by Road Surface Condition
```{r}


PLCD$Crash_Date_YMD<- ymd(substr(PLCD$Crash_Date,1,10))
months(PLCD$Crash_Date_YMD)

PLCD$Crash_Date_YM<- format(PLCD$Crash_Date_YMD,"%Y %B")


RSC <- as.data.frame(
  PLCD%>%
    filter(!is.na(Crash_Date_YM)) %>%
    group_by(Crash_Date_YMD,Road_Surface_Condition) %>%
    summarise(Number_of_Crashes = n())
                     )

#write.csv(RSC,"RSC.csv")



ggplot(RSC,aes(x=RSC$Crash_Date_YMD,y=RSC$Number_of_Crashes))+geom_line(aes(color=Road_Surface_Condition),alpha=0.8)+theme(axis.text.x = element_text(angle = 90, hjust = 1))

#dev.new()
```



```{r}

unique(PLCD$Traffic_Control_Device_Functioning)

year(PLCD$Crash_Date_YMD)


PLCD%>%
  group_by(County_Name)%>%
  filter(Traffic_Control_Device_Fnuctioning=="No, device not functioning" & year(Crash_Date_YMD)==2019)%>%
  summarise(ct =n())

PLCD%>%
  group_by(County_Name)%>%
  filter(Traffic_Control_Device_Functioning=="No, device not functioning" & year(Crash_Date_YMD)==2019)%>%
  summarise(ct = n())

```



```{r}

#Import County Wise Population
County_Pop <- read.csv("County_Year_Pop_Mass.csv")
County_Pop <- gather(County_Pop, key = "Year", value = "Population" ,  2:5)
County_Pop$Year<- as.numeric(str_replace(County_Year_Pop_Mass$Year,"X",""))

#importing 2019 population
Mass_Counties_Pop_2019 <- read.csv("Mass_Counties_Pop_2019.csv")
County_Pop<-rbind(County_Pop,Mass_Counties_Pop_2019)
rm(Mass_Counties_Pop_2019)


############################
#Total Number of Crashes
score_df<-PLCD%>%
  group_by(County_Name,Crash_Year)%>%
  summarise(Number_of_Crashes =n())

colnames(score_df)[colnames(score_df)=="County_Name"] <- "County"
colnames(score_df)[colnames(score_df)=="Crash_Year"] <- "Year"
Crash_Percentage_Vs_Pop <- inner_join(County_Pop,score_df)

Crash_Percentage_Vs_Pop$Crash_Percentage<-Crash_Percentage_Vs_Pop$Number_of_Crashes/Crash_Percentage_Vs_Pop$Population

County_Crash_Percentage <-
  Crash_Percentage_Vs_Pop %>%
  group_by(County) %>%
  summarise(Crash_Percentage = mean(Crash_Percentage))


#median(County_Crash_Percentage$Crash_Percentage)


County_Crash_Percentage$Mean_Crash_Percent <- mean(County_Crash_Percentage$Crash_Percentage)

County_Crash_Percentage$Crash_Score <- County_Crash_Percentage$Crash_Percentage/County_Crash_Percentage$Mean_Crash_Percent




#Traffic Device Functioning
TCDNF_score_df<-PLCD%>%
  filter(Traffic_Control_Device_Functioning == "No, device not functioning")%>%
  group_by(County_Name,Crash_Year)%>%
  summarise(TCDNF_Number_of_Crashes =n())


colnames(TCDNF_score_df)[colnames(TCDNF_score_df)=="County_Name"] <- "County"
colnames(TCDNF_score_df)[colnames(TCDNF_score_df)=="Crash_Year"] <- "Year"
Crash_Percentage_Vs_Pop <- inner_join(County_Pop,TCDNF_score_df)

#checking PLCD for NA in Crash_Year
PLCD%>%
  filter(is.na(Crash_Year))

Crash_Percentage_Vs_Pop$Crash_Percentage<-Crash_Percentage_Vs_Pop$Number_of_Crashes/Crash_Percentage_Vs_Pop$Population

County_Crash_Percentage <-
  Crash_Percentage_Vs_Pop %>%
  group_by(County) %>%
  summarise(Crash_Percentage = mean(Crash_Percentage))


#median(County_Crash_Percentage$Crash_Percentage)


County_Crash_Percentage$Mean_Crash_Percent <- mean(County_Crash_Percentage$Crash_Percentage)

County_Crash_Percentage$Crash_Score <- County_Crash_Percentage$Crash_Percentage/County_Crash_Percentage$Mean_Crash_Percent


#Protective System Used

#Alcohol Suspected

#Drugs Suspected




```

